Decoupled Fallback: A Portable Single-Pass GPU Scan

### 2025-02-26 paper.tex
linted (actually, chktex)

\subsubsection{Fairness and Progress}
Because multiple workgroups may reside on a single multiprocessor, two key issues arise at the subgroup scheduling level: \emph{fairness}---how evenly execution resources are distributed among subgroups---and \emph{progress guarantees}---ensuring that subgroups eventually make progress towards termination. However, the term "fairness" is used differently across the literature, leading to potential confusion. Sorensen et al.~\cite{sorensen2016,sorensen2018,sorensen2021}, whose work provides the most comprehensive taxonomy of progress models to date, use "fairness" specifically to describe differing levels of progress guarantees. In contrast, NVIDIA publications~\cite{4523358,Merrill2016} define "fairness" in terms of the distribution of execution resources among subgroups. In this work, we adopt NVIDIA's terminology.

@@

\subsubsection{Scan-then-Propagate}
Introduced by Sengupta et al.~\cite{10.5555/1280094.1280110}, \emph{Scan-then-Propagate}~\cite{GPUGems3, Sengupta2011} extends intra-workgroup scans to large inputs by resolving inter-workgroup dependencies in three phases:
\begin{enumerate}
  \item \textbf{Intra-Workgroup (Local) Scan:}  Each workgroup performs an inclusive scan on its assigned work tile, storing local results and reductions in global memory.
  \item \textbf{Spine Scan:} The workgroup reductions, collectively forming the \emph{spine}, are gathered and scanned to compute the \emph{root}, resolving dependencies between work tiles. For large scans, this stage is an insignificant amount of work compared to the first and third.
  \item \textbf{Propagation:} Each workgroup retrieves its portion of the root and adds it to the local scan output.
\end{enumerate}
The absence of inter-workgroup barriers necessitates separate kernel launches and pipeline barriers for dependency resolution, naturally enforcing this structure. To accommodate arbitrarily large inputs, Sengupta et al. employ recursion, repeatedly applying the three-phase process until the spine fits within a single work tile. Thus, an input of size $n$ and a work tile of size $t$ results in a recursive depth of $\lceil \log_t n \rceil$, and $2\cdot\lceil \log_t n \rceil - 1$ kernel launches. Each recursive step $k$ processes $n/t^k$ tiles, leading to a total of $\sum_{k=1}^{\lceil \log_t n \rceil} n/t^k \approx (n - 1)/(t - 1)$ work tiles over all steps. Since each tile moves $4t$ data ($2t$ local scan, $2t$ propagation), the total global data movement is $O\left(\frac{4t(n - 1)}{t - 1}\right) = O(4n)$.

@@

\subsection{Why does \emph{Chained-Scan} Rely on Forward-Progress Guarantees?}
A \emph{forward-progress guarantee} (FPG) ensures that all occupant subgroups (of possibly different workgroups) eventually make progress toward termination, preventing indefinite starvation. Prior to Sorensen et al.~\cite{sorensen2018,sorensen2021}, developers largely relied on untested, folklore-like assumptions about GPU progress guarantees. Instead, Sorensen et al. systematically defined and tested three progress models\footnote{Note, a ``workgroup'' in Sorensen et al.'s definition is not necessarily in-flight and occupant on a multiprocessor.}:
\begin{itemize}
  \item \textbf{Occupancy-Bound Execution (OBE)}: Workgroups that have made at least one execution step.
  \item \textbf{Heterogeneous System Architecture (HSA)}: The workgroup with the lowest ID that has not yet terminated.
  \item \textbf{Linear OBE (LOBE)}: Workgroups that have made at least one step, and any workgroup with a lower ID than a workgroup that has made at least one step.
\end{itemize}
Their findings demonstrated that Apple and ARM GPUs do not support LOBE, contradicting their prior assumptions in \cite{sorensen2018}.

@@

\begin{table}
  \small
  \centering
  \begin{tabular}{lcccc}
    \toprule
    Algorithm       & LOBE & OBE & HSA & Unfair \\
    \midrule
    CS (Workgroup ID)  & Yes & No  & Yes & No  \\
    CS (Atomic ID)        & Yes & Yes & No  & No  \\
    Ours               & Yes & Yes & Yes & Yes \\
    \bottomrule
  \end{tabular}
  \caption{Comparison of Chained Scan (CS) Variants and Our Implementation Across Progress Models.}
  \label{tab:progress_models}
\end{table}

@@

The viability of \emph{Decoupled-Fallback} relies on the same cache locality properties as \emph{Decoupled-Lookback}. Since no subgroup scheduler is perfectly fair, all hardware inherently has a natural fallback rate $f$\!, independent of FPG, due to false-positive fallback operations. As a result, memory bandwidth usage increases (and we expect throughput to decrease) by $O(fn)$ to $O((2 + f)n)$. However, just as serial ordering confines spine memory accesses within an $o$-width sliding window, scan input accesses are likewise constrained to an $ot$-width window, increasing their likelihood of residing in cache. As we later empirically demonstrate in \ref{sec:Sim-Blocking}, this cache residency significantly reduces memory access costs, making fallback operations predominantly compute-bound when reducing a blocking tile. Since pure reduction has minimal overhead, throughput approaches $O(2n)$ on most devices.

@@

        \If{$\textit{cond1} \textbf{ \&\& } \textit{cond2}$}{
          $\textit{x}[\textit{fanout\_index}] \gets \textit{x}[\textit{fanout\_index}] + \textit{x}[((\textit{fanout\_index} / \textit{stride}) + \textit{top\_offset}) - 1]$\;
        }
      }
      $\textit{stride} \gets \textit{stride} \cdot \textit{S}$\;
      $\textit{top\_offset} \gets \textit{top\_offset} + \textit{step}$\;
    }
    \textbf{barrier()}\;
  }
  \Return{$\textit{x}$}\;
  \caption{Workgroup-Wide Scan.}
  \label{alg:example}
\end{algorithm}

@@

We now detail the \emph{Decoupled-Fallback} procedure, presenting the algorithm from the perspective of a single workgroup. The details of tile state encoding and atomic updates are discussed in the following section and may vary depending on the data type, so for simplicity, we describe the operation as being performed by the \emph{first thread} of the workgroup:
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{graphics/FlowChart.pdf}
  \label{fig:decoupled_fallback}
  \caption{Flow Chart Representation of Decoupled Fallback}
\end{figure}
\begin{enumerate}
  \item[(0)] \textbf{Posting Local Reduction:}
        \begin{itemize}
          \item Each workgroup posts its local reduction---the result of the workgroup-wide scan over partials---to global memory and updates its tile state to \emph{Ready}.
          \item If it is the first work tile, its local reduction \emph{is} its inclusive reduction. In this case, the workgroup updates its tile state directly to \emph{Inclusive} and bypasses the \emph{Chained-Propagation} phase entirely.
        \end{itemize}

@@

\section{Evaluation}
\subsection{Experimental Setup}
We configure workgroups with a size $w = 256$ threads, where each thread processes $p = 4$ vectors of size 4, resulting in 16 elements per thread and a total work tile size of $t = 4096$. We find a maximum spin count of 4 effectively balances responsiveness to blocking while minimizing false-positive fallbacks. As the size of the local spine is inversely proportional to the subgroup size, we allocate sufficient shared memory to accommodate the WebGPU standard's minimum subgroup size of 4. Due to our overall thriftiness with shared memory, total usage remains modest at only $\sim$1 KB\@. For all tests, we use an input size of $2^{25}$ and an inclusive prefix sum. Our test hardware, listed POSITIONAL-RELATION, spans a wide range of FPG and performance capabilities. We empirically find that Google Chrome's native implementation of WebGPU, Dawn, provides the best performance and use it for all testing, except on the Arm-Mali GPU, which requires Android native, necessitating the use of WGPU\@.\footnote{To keep the perforamance as close to Dawn as possible, we translate our WGSL code with Dawn's Tint transpiler, and provide the raw SPIR-V to WGPU\@.}
\begin{table}
  \centering
  \small
  \setlength{\tabcolsep}{1pt}
  \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} l l c r r}
    \toprule
    Vendor & Device Name & LOBE & Occupancy & Subgroup Size \\
    \midrule
    Intel  & HD Graphics 620        & \textbf{Yes} & 9   & 8--32  \\
    ARM    & Mali-G78 MP20          & \textbf{No}  & 40  & 16  \\
    Apple  & M3     (10 Cores)      & \textbf{No}  & 80  & 32  \\
    Apple  & M1 Max (32 Cores)      & \textbf{No}  & 160 & 32  \\
    NVIDIA & GeForce RTX 2080 Super & \textbf{Yes} & 192 & 32  \\
    AMD    & Radeon RX 7900 XT      & \textbf{Yes} & 420 & 32--64  \\
    \bottomrule
  \end{tabular*}
  \caption{Hardware Specifications of Testing Devices.}
  \label{tab:hardware}
\end{table}

@@

Cross-API comparisons remain challenging, as our CUDA implementation consistently outperforms its Dawn counterpart despite near-identical high-level implementations. The performance cost of \emph{DF} is most pronounced in CUDA when the entire input fits within the 2080 Super’s 4MB L2 cache, where reduced global memory latency exposes \emph{DF's} increased synchronization and compute overhead compared to CUB. However, at larger input sizes, where memory accesses exceed cache capacity, \emph{DF} reaches parity with CUB and achieves \emph{Memcpy} speeds.

@@

\begin{table*}
  \centering
  \setlength{\tabcolsep}{1pt}
  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l r r r r r r}
    \toprule
    Metric & Intel HD620 & ARM Mali-G78 MP20 & Apple M3 & Apple M1 Max & Nvidia 2080 Super & AMD 7900 XT \\
    \midrule
    Avg \emph{Memcpy} (32-bit ele/sec $\cdot 10^9$)  & 1.50   & 2.48  & 10.25  & 38.62  & 54.39  & 72.28  \\
    Avg \emph{DF} (32-bit ele/sec $\cdot 10^9$)   & 1.49   & 2.79  & 10.22  & 38.84  & 53.29  & 94.60  \\
    Avg \emph{RTS} (32-bit ele/sec $\cdot 10^9$)      & 1.06   & 2.02  & 6.81   & 25.13  & 35.90  & 61.49  \\
    \emph{DF}/\emph{Memcpy}       & 99.1\%    & 112.8\%  & 99.8\%    & 100.6\%   & 98.0\%    & 130.9\%   \\
    \emph{DF}/\emph{RTS}          & 140.2\%   & 138.5\%  & 150.2\%   & 154.5\%   & 148.4\%   & 153.8\%   \\
    Avg Spins per Wg    & 0.305    & 1.477    & 0.606    & 0.604    & 0.419    & 0.148    \\
    Avg Lookback Length per Wg  & 1.726    & 3.537    & 3.415    & 10.782   & 13.046   & 20.722   \\
    Avg Total Fallbacks Initiated      & 254.662   & 2504.704  & 365.546   & 990.962   & 86.400   & 9.308   \\
    Avg Total Successful Insertions    & 6.170     & 122.436   & 4.612     & 0.012    & 0.284    & 0.020    \\
    \bottomrule
  \end{tabular*}
  \caption{\emph{Decoupled Fallback} performance, 32-bit Inclusive Prefix Sum, $\mathbf{2^{25}}$ 32-bit elements}
  \label{tab:results}
\end{table*}

@@

\subsection{Simulated Blocking}\label{sec:Sim-Blocking}
To demonstrate the resilience of our technique to unfair schedulers, we simulate fallback rates significantly higher than those observed organically by intentionally omitting work tile reduction updates. Because these updates are skipped, the immediate successor workgroup reads \emph{Not Ready}, spinning and triggering a fallback. Starting with infrequent omissions---1 in 512 work tiles---we progressively increase the omission rate to near-continuous levels---1 in 2 work tiles----recording \emph{DF} statistics at each step.
\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/fallbacksInitiated_plot.pdf}
    \caption{Fallbacks Initiated}
    \label{fig:fallbacks_initiated}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/totalSpins_plot.pdf}
    \caption{Total Spins}
    \label{fig:total_spins}
  \end{subfigure}

@@

  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/lookbackLength_plot.pdf}
    \caption{Lookback Length}
    \label{fig:lookback_length}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/time_plot.pdf}
    \caption{Throughput vs.\ Baseline}
    \label{fig:execution_time}
  \end{subfigure}

  \caption{Simulated Blocking. \john{Please write a good caption here---this is an important caption to write! Probably writing subfigure captions is the right way to go (explain each of them individually).}}
  \label{fig:vertical_images}
\end{figure*}

As shown in Figure~\ref{fig:execution_time}, performance relative to a no-omission baseline remains above 85\% even at near 100\% fallback rates across all devices---except for the Intel HD620, where only a 25\% fallback rate is induced. Notably, this performance level remains well above the equivalent \emph{RTS} performance of 66\%. While Figures~\ref{fig:fallbacks_initiated} and \ref{fig:total_spins} confirm that we are inducing fallbacks, this test primarily measures the impact of increasing fallback operations rather than true blocking behavior. The finer details of this distinction are discussed next.
\section{Discussion}

@@

Our reexamination of \emph{Chained Scan} with dynamic assignment raises fundamental questions about the GPU progress models proprosed by Sorensen et al. in \cite{sorensen2021}: Can all HSA algorithms be converted to OBE via dynamic assignment? Are there cases, excluding atomic memory errors, where LOBE guarantees termination but dynamic assignment with OBE does not? Is OBE all you need?

@@

\subsection{Baseline Comparisons: \emph{Memcpy}}
\label{sec:memcpy}
\begin{table}
  \small
  \centering
  \begin{tabular}{l r r}
    \toprule
    Device                                & \makecell{Memory                   \\ Bandwidth \\ (GB/s)} & \makecell{Expected \\ Throughput \\ (32-bit ele/sec)} \\
    \midrule
    Intel HD620 Single-Channel            & 17.1             & 2.13e9          \\
    Intel HD620 Dual-Channel              & 34.1             & 4.26e9          \\
    ARM Mali-G78 MP20 / Tensor (Pixel 6a) & 51.2             & 6.40e9          \\
    Apple M3 (10 Cores)                   & 102.4            & 12.8e9          \\
    Apple M1 Max (32 Cores)               & 409.6            & 51.2e9          \\
    Nvidia 2080 Super                     & 496              & 62.0e9          \\
    AMD 7900 XT                           & 800              & 100.0e9         \\
    \bottomrule
  \end{tabular}
  \caption{Vendor-Specified Memory Bandwidth and Expected Throughput (32-bit Elements per Second).}
  \label{tab:memory_bandwidth}
\end{table}
Although all of Dawn's backend APIs support \emph{timestamp queries inside encoders}, Apple GPUs do not, so we are forced to implement our own \emph{Memcpy} kernel to measure memory bandwidth, as opposed to simply timing an API's copy buffer operation. It is well known that \emph{observed} peak memory bandwidth, tends to diverge from vendor-listed peak bandwidth, so for completeness we list these figures above. We make a number of observations:
\begin{itemize}
  \item \textbf{Intel HD620:} As our HD620 laptop is only equipped with a single memory DIMM, its peak bandwidth is limited to the lower bound, 2.13e9 ele/sec. Thus, our observed bandwidth of 1.498e9 is reasonable.
  \item \textbf{ARM Mali-G78 MP20 / Pixel 6a (Tensor):} The observed throughput is significantly lower than expected, despite a vendor-listed bandwidth of 6.40e9 ele/sec. The cause of this discrepancy remains unclear, but we note other developers have observed similar speeds.
  \item \textbf{Apple M1 Pro Max:} Observed throughput matches expectations.
  \item \textbf{Nvidia 2080 Super:} Observed throughput matches expectations.
  \item \textbf{AMD 7900 XT:} Our above result, where \emph{DF} was observed to be faster than \emph{Memcopy}, is likely due to compute latency counterintuitively improving memory access patterns, leading to reduced paging. Furthermore, we see that our \emph{DF} throughput is within spec and reasonable.
\end{itemize}

\subsection{Baseline Comparisons: \emph{Reduce then Scan}}
\label{sec:rts}
As we desire the most competitve baseline possible, our \emph{RTS} kernels employ the same intra-workgroup scan strategy and tuning parameters as \emph{DF}. Initially, we adopted Merrill and Grimshaw's~\cite{Merrill2009} workgroup raking approach for inter-workgroup processing. However, empirical testing revealed that raking was consistently outperformed—by up to $\sim$15\% on lower-end devices—by a simpler approach that processes the spine serially with a single workgroup. Furthermore, our \emph{RTS} performance closely aligns with the theoretically expected $O(3n)$ global memory movement.

>>>>

\subsubsection{Fairness and Progress}
Because multiple workgroups may reside on a single multiprocessor, two key issues arise at the subgroup scheduling level: \emph{fairness}---how evenly execution resources are distributed among subgroups---and \emph{progress guarantees}---ensuring that subgroups eventually make progress towards termination. However, the term ``fairness'' is used differently across the literature, leading to potential confusion. Sorensen et al.~\cite{sorensen2016,sorensen2018,sorensen2021}, whose work provides the most comprehensive taxonomy of progress models to date, use ``fairness'' specifically to describe differing levels of progress guarantees. In contrast, NVIDIA publications~\cite{4523358,Merrill2016} define ``fairness'' in terms of the distribution of execution resources among subgroups. In this work, we adopt NVIDIA's terminology.

@@

\subsubsection{Scan-then-Propagate}
Introduced by Sengupta et al.~\cite{10.5555/1280094.1280110}, \emph{Scan-then-Propagate}~\cite{GPUGems3, Sengupta2011} extends intra-workgroup scans to large inputs by resolving inter-workgroup dependencies in three phases:
\begin{enumerate}
  \item \textbf{Intra-Workgroup (Local) Scan:}  Each workgroup performs an inclusive scan on its assigned work tile, storing local results and reductions in global memory.
  \item \textbf{Spine Scan:} The workgroup reductions, collectively forming the \emph{spine}, are gathered and scanned to compute the \emph{root}, resolving dependencies between work tiles. For large scans, this stage is an insignificant amount of work compared to the first and third.
  \item \textbf{Propagation:} Each workgroup retrieves its portion of the root and adds it to the local scan output.
\end{enumerate}
The absence of inter-workgroup barriers necessitates separate kernel launches and pipeline barriers for dependency resolution, naturally enforcing this structure. To accommodate arbitrarily large inputs, Sengupta et al.\ employ recursion, repeatedly applying the three-phase process until the spine fits within a single work tile. Thus, an input of size $n$ and a work tile of size $t$ results in a recursive depth of $\lceil \log_t n \rceil$, and $2\cdot\lceil \log_t n \rceil - 1$ kernel launches. Each recursive step $k$ processes $n/t^k$ tiles, leading to a total of $\sum_{k=1}^{\lceil \log_t n \rceil} n/t^k \approx (n - 1)/(t - 1)$ work tiles over all steps. Since each tile moves $4t$ data ($2t$ local scan, $2t$ propagation), the total global data movement is $O\left(\frac{4t(n - 1)}{t - 1}\right) = O(4n)$.

@@

\subsection{Why does \emph{Chained-Scan} Rely on Forward-Progress Guarantees?}
A \emph{forward-progress guarantee} (FPG) ensures that all occupant subgroups (of possibly different workgroups) eventually make progress toward termination, preventing indefinite starvation. Prior to Sorensen et al.~\cite{sorensen2018,sorensen2021}, developers largely relied on untested, folklore-like assumptions about GPU progress guarantees. Instead, Sorensen et al.\ systematically defined and tested three progress models:\footnote{Note, a ``workgroup'' in Sorensen et al.'s definition is not necessarily in-flight and occupant on a multiprocessor.}
\begin{itemize}
  \item \textbf{Occupancy-Bound Execution (OBE)}: Workgroups that have made at least one execution step.
  \item \textbf{Heterogeneous System Architecture (HSA)}: The workgroup with the lowest ID that has not yet terminated.
  \item \textbf{Linear OBE (LOBE)}: Workgroups that have made at least one step, and any workgroup with a lower ID than a workgroup that has made at least one step.
\end{itemize}
Their findings demonstrated that Apple and ARM GPUs do not support LOBE, contradicting their prior assumptions~\cite{sorensen2018}.

@@

\begin{table}
  \small
  \centering
  \begin{tabular}{lcccc}
    \toprule
    Algorithm       & LOBE & OBE & HSA & Unfair \\
    \midrule
    CS (Workgroup ID)  & Yes & No  & Yes & No  \\
    CS (Atomic ID)        & Yes & Yes & No  & No  \\
    Ours               & Yes & Yes & Yes & Yes \\
    \bottomrule
  \end{tabular}
  \caption{Comparison of Chained Scan (CS) Variants and Our Implementation Across Progress Models.\label{tab:progress_models}}
\end{table}

@@

The viability of \emph{Decoupled-Fallback} relies on the same cache locality properties as \emph{Decoupled-Lookback}. Since no subgroup scheduler is perfectly fair, all hardware inherently has a natural fallback rate $f$\!, independent of FPG, due to false-positive fallback operations. As a result, memory bandwidth usage increases (and we expect throughput to decrease) by $O(fn)$ to $O((2 + f)n)$. However, just as serial ordering confines spine memory accesses within an $o$-width sliding window, scan input accesses are likewise constrained to an $ot$-width window, increasing their likelihood of residing in cache. As we later empirically demonstrate in Section~\ref{sec:Sim-Blocking}, this cache residency significantly reduces memory access costs, making fallback operations predominantly compute-bound when reducing a blocking tile. Since pure reduction has minimal overhead, throughput approaches $O(2n)$ on most devices.

@@

        \If{$\textit{cond1} \textbf{ \&\& } \textit{cond2}$}{
          $\textit{x}[\textit{fanout\_index}] \gets \textit{x}[\textit{fanout\_index}] + \textit{x}[((\textit{fanout\_index} / \textit{stride}) + \textit{top\_offset}) - 1]$\;
        }
      }
      $\textit{stride} \gets \textit{stride} \cdot \textit{S}$\;
      $\textit{top\_offset} \gets \textit{top\_offset} + \textit{step}$\;
    }
    \textbf{barrier()}\;
  }
  \Return{$\textit{x}$}\;
  \caption{Workgroup-Wide Scan.\label{alg:example}}
\end{algorithm}

@@

We now detail the \emph{Decoupled-Fallback} procedure, presenting the algorithm from the perspective of a single workgroup. The details of tile state encoding and atomic updates are discussed in the following section and may vary depending on the data type, so for simplicity, we describe the operation as being performed by the \emph{first thread} of the workgroup:
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{graphics/FlowChart.pdf}
  \caption{Flow Chart Representation of Decoupled Fallback.\label{fig:decoupled_fallback}}
\end{figure}
\begin{enumerate}
  \item[(0)] \textbf{Posting Local Reduction:}
        \begin{itemize}
          \item Each workgroup posts its local reduction---the result of the workgroup-wide scan over partials---to global memory and updates its tile state to \emph{Ready}.
          \item If it is the first work tile, its local reduction \emph{is} its inclusive reduction. In this case, the workgroup updates its tile state directly to \emph{Inclusive} and bypasses the \emph{Chained-Propagation} phase entirely.
        \end{itemize}

@@

\section{Evaluation}
\subsection{Experimental Setup}
We configure workgroups with a size $w = 256$ threads, where each thread processes $p = 4$ vectors of size 4, resulting in 16 elements per thread and a total work tile size of $t = 4096$. We find a maximum spin count of 4 effectively balances responsiveness to blocking while minimizing false-positive fallbacks. As the size of the local spine is inversely proportional to the subgroup size, we allocate sufficient shared memory to accommodate the WebGPU standard's minimum subgroup size of 4. Due to our overall thriftiness with shared memory, total usage remains modest at only $\sim$1 KB\@. For all tests, we use an input size of $2^{25}$ and an inclusive prefix sum. Our test hardware, listed POSITIONAL-RELATION, spans a wide range of FPG and performance capabilities. We empirically find that Google Chrome's native implementation of WebGPU, Dawn, provides the best performance and use it for all testing, except on the Arm-Mali GPU, which requires Android native, necessitating the use of WGPU\@.\footnote{To keep the perforamance as close to Dawn as possible, we translate our WGSL code with Dawn's Tint transpiler, and provide the raw SPIR-V to WGPU\@.}
\begin{table}
  \centering
  \small
  \setlength{\tabcolsep}{1pt}
  \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} l l c r r}
    \toprule
    Vendor & Device Name & LOBE & Occupancy & Subgroup Size \\
    \midrule
    Intel  & HD Graphics 620        & \textbf{Yes} & 9   & 8--32  \\
    ARM    & Mali-G78 MP20          & \textbf{No}  & 40  & 16  \\
    Apple  & M3     (10 Cores)      & \textbf{No}  & 80  & 32  \\
    Apple  & M1 Max (32 Cores)      & \textbf{No}  & 160 & 32  \\
    NVIDIA & GeForce RTX 2080 Super & \textbf{Yes} & 192 & 32  \\
    AMD    & Radeon RX 7900 XT      & \textbf{Yes} & 420 & 32--64  \\
    \bottomrule
  \end{tabular*}
  \caption{Hardware Specifications of Testing Devices.\label{tab:hardware}}
\end{table}

@@

Cross-API comparisons remain challenging, as our CUDA implementation consistently outperforms its Dawn counterpart despite near-identical high-level implementations. The performance cost of \emph{DF} is most pronounced in CUDA when the entire input fits within the 2080 Super’s 4MB L2 cache, where reduced global memory latency exposes \emph{DF's} increased synchronization and compute overhead compared to CUB\@. However, at larger input sizes, where memory accesses exceed cache capacity, \emph{DF} reaches parity with CUB and achieves \emph{Memcpy} speeds.

@@

\begin{table*}
  \centering
  \setlength{\tabcolsep}{1pt}
  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l r r r r r r}
    \toprule
    Metric & Intel HD620 & ARM Mali-G78 MP20 & Apple M3 & Apple M1 Max & Nvidia 2080 Super & AMD 7900 XT \\
    \midrule
    Avg \emph{Memcpy} (32-bit ele/sec $\cdot 10^9$)  & 1.50   & 2.48  & 10.25  & 38.62  & 54.39  & 72.28  \\
    Avg \emph{DF} (32-bit ele/sec $\cdot 10^9$)   & 1.49   & 2.79  & 10.22  & 38.84  & 53.29  & 94.60  \\
    Avg \emph{RTS} (32-bit ele/sec $\cdot 10^9$)      & 1.06   & 2.02  & 6.81   & 25.13  & 35.90  & 61.49  \\
    \emph{DF}/\emph{Memcpy}       & 99.1\%    & 112.8\%  & 99.8\%    & 100.6\%   & 98.0\%    & 130.9\%   \\
    \emph{DF}/\emph{RTS}          & 140.2\%   & 138.5\%  & 150.2\%   & 154.5\%   & 148.4\%   & 153.8\%   \\
    Avg Spins per Wg    & 0.305    & 1.477    & 0.606    & 0.604    & 0.419    & 0.148    \\
    Avg Lookback Length per Wg  & 1.726    & 3.537    & 3.415    & 10.782   & 13.046   & 20.722   \\
    Avg Total Fallbacks Initiated      & 254.662   & 2504.704  & 365.546   & 990.962   & 86.400   & 9.308   \\
    Avg Total Successful Insertions    & 6.170     & 122.436   & 4.612     & 0.012    & 0.284    & 0.020    \\
    \bottomrule
  \end{tabular*}
  \caption{\emph{Decoupled Fallback} performance, 32-bit Inclusive Prefix Sum, $\mathbf{2^{25}}$ 32-bit elements.\label{tab:results}}
\end{table*}

@@

\subsection{Simulated Blocking}\label{sec:Sim-Blocking}
To demonstrate the resilience of our technique to unfair schedulers, we simulate fallback rates significantly higher than those observed organically by intentionally omitting work tile reduction updates. Because these updates are skipped, the immediate successor workgroup reads \emph{Not Ready}, spinning and triggering a fallback. Starting with infrequent omissions---1 in 512 work tiles---we progressively increase the omission rate to near-continuous levels---1 in 2 work tiles---recording \emph{DF} statistics at each step.
\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/fallbacksInitiated_plot.pdf}
    \caption{Fallbacks Initiated.\label{fig:fallbacks_initiated}}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/totalSpins_plot.pdf}
    \caption{Total Spins.\label{fig:total_spins}}
  \end{subfigure}

@@

  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/lookbackLength_plot.pdf}
    \caption{Lookback Length.\label{fig:lookback_length}}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/time_plot.pdf}
    \caption{Throughput vs.\ Baseline.\label{fig:execution_time}}
  \end{subfigure}

  \caption{Simulated Blocking. \john{Please write a good caption here---this is an important caption to write! Probably writing subfigure captions is the right way to go (explain each of them individually).}\label{fig:vertical_images}}
\end{figure*}

As shown in Figure~\ref{fig:execution_time}, performance relative to a no-omission baseline remains above 85\% even at near 100\% fallback rates across all devices---except for the Intel HD620, where only a 25\% fallback rate is induced. Notably, this performance level remains well above the equivalent \emph{RTS} performance of 66\%. While Figures~\ref{fig:fallbacks_initiated} and~\ref{fig:total_spins} confirm that we are inducing fallbacks, this test primarily measures the impact of increasing fallback operations rather than true blocking behavior. The finer details of this distinction are discussed next.
\section{Discussion}

@@

Our reexamination of \emph{Chained Scan} with dynamic assignment raises fundamental questions about the GPU progress models proprosed by Sorensen et al.~\cite{sorensen2021}: Can all HSA algorithms be converted to OBE via dynamic assignment? Are there cases, excluding atomic memory errors, where LOBE guarantees termination but dynamic assignment with OBE does not? Is OBE all you need?

@@

\subsection{Baseline Comparisons: \emph{Memcpy}}%
\label{sec:memcpy}
\begin{table}
  \small
  \centering
  \begin{tabular}{l r r}
    \toprule
    Device                                & \makecell{Memory                   \\ Bandwidth \\ (GB/s)} & \makecell{Expected \\ Throughput \\ (32-bit ele/sec)} \\
    \midrule
    Intel HD620 Single-Channel            & 17.1             & 2.13e9          \\
    Intel HD620 Dual-Channel              & 34.1             & 4.26e9          \\
    ARM Mali-G78 MP20 / Tensor (Pixel 6a) & 51.2             & 6.40e9          \\
    Apple M3 (10 Cores)                   & 102.4            & 12.8e9          \\
    Apple M1 Max (32 Cores)               & 409.6            & 51.2e9          \\
    Nvidia 2080 Super                     & 496              & 62.0e9          \\
    AMD 7900 XT                           & 800              & 100.0e9         \\
    \bottomrule
  \end{tabular}
  \caption{Vendor-Specified Memory Bandwidth and Expected Throughput (32-bit Elements per Second).\label{tab:memory_bandwidth}}
\end{table}
Although all of Dawn's backend APIs support \emph{timestamp queries inside encoders}, Apple GPUs do not, so we are forced to implement our own \emph{Memcpy} kernel to measure memory bandwidth, as opposed to simply timing an API's copy buffer operation. It is well known that \emph{observed} peak memory bandwidth, tends to diverge from vendor-listed peak bandwidth, so for completeness we list these figures above. We make a number of observations:
\begin{itemize}
  \item \textbf{Intel HD620:} As our HD620 laptop is only equipped with a single memory DIMM, its peak bandwidth is limited to the lower bound, 2.13e9 ele/sec. Thus, our observed bandwidth of 1.498e9 is reasonable.
  \item \textbf{ARM Mali-G78 MP20 / Pixel 6a (Tensor):} The observed throughput is significantly lower than expected, despite a vendor-listed bandwidth of 6.40e9 ele/sec. The cause of this discrepancy remains unclear, but we note other developers have observed similar speeds.
  \item \textbf{Apple M1 Pro Max:} Observed throughput matches expectations.
  \item \textbf{Nvidia 2080 Super:} Observed throughput matches expectations.
  \item \textbf{AMD 7900 XT:} Our above result, where \emph{DF} was observed to be faster than \emph{Memcopy}, is likely due to compute latency counterintuitively improving memory access patterns, leading to reduced paging. Furthermore, we see that our \emph{DF} throughput is within spec and reasonable.
\end{itemize}

\subsection{Baseline Comparisons: \emph{Reduce then Scan}}%
\label{sec:rts}
As we desire the most competitve baseline possible, our \emph{RTS} kernels employ the same intra-workgroup scan strategy and tuning parameters as \emph{DF}. Initially, we adopted Merrill and Grimshaw's~\cite{Merrill2009} workgroup raking approach for inter-workgroup processing. However, empirical testing revealed that raking was consistently outperformed—by up to $\sim$15\% on lower-end devices—by a simpler approach that processes the spine serially with a single workgroup. Furthermore, our \emph{RTS} performance closely aligns with the theoretically expected $O(3n)$ global memory movement.
